\propertitle

### 2019-04-09 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-04-08T13:52:04+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation constraints%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample constraints}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\renewcommand*{\cites}{\parencites}

@@

@@

To apply the \me\ method to the larger, unsampled population, it is
necessary to use probability relations relevant to sampling
\cites{ghoshetal1997}[parts~I,
VI]{freedmanetal1978_r2007}[\chap~8]{gelmanetal1995_r2014}[\chap~3]{jaynes1994_r2003}.
The relations we present are well-known in survey sampling and in the
pedagogic problem of drawing from an urn without replacement, yet they are
somewhat hard to find explicitly written in the neuroscientific literature.
We present and discuss them in the next section. A minor purpose of this
paper is to make these relations more widely known, because they can be
useful independently of \me\ methods.

@@

Our uncertainties about the full population and the sample are connected
via the conditional probability
\begin{equation}
  \label{eq:conditional_hypergeometric}
  \pf(\yrs \|\yRf, \yH)=
  \binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}
  \defs \ypp_{\yrs\yRf},
\end{equation}
which is a hypergeometric distribution, typical of \enquote{drawing without
  replacement} problems. The combinatorial proof of this expression is in
fact the same as for this class of problems
\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.

@@

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-04-09T17:16:14+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation constraints%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample constraints}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

%\newcommand*{\citeps}{\parencites}

@@

\newcommand*{\ynu}{\bm{\nu}}

@@

To apply the \me\ method to the larger, unsampled population, it is
necessary to use probability relations relevant to sampling
\citep{ghoshetal1997}[parts~I,
VI]{freedmanetal1978_r2007}[\chap~8]{gelmanetal1995_r2014}[\chap~3]{jaynes1994_r2003}.
The relations we present are well-known in survey sampling and in the
pedagogic problem of drawing from an urn without replacement, yet they are
somewhat hard to find explicitly written in the neuroscientific literature.
We present and discuss them in the next section. A minor purpose of this
paper is to make these relations more widely known, because they can be
useful independently of \me\ methods.

@@

Our uncertainties about the full population and the sample are connected
via the conditional probability
\begin{equation}
  \label{eq:conditional_hypergeometric}
  \pf(\yrs \|\yRf, \yH)=
  \binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}
  \defs \ypp_{\yrs\yRf},
\end{equation}
which is a hypergeometric distribution, typical of \enquote{drawing without
  replacement} problems. The combinatorial proof of this expression is in
fact the same as for this class of problems
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.

@@

There are three inequivalent main routes that lead to a probability
distribution of the maximum-entropy form~\eqref{eq:app_maxent_sample}
or~\eqref{eq:app_maxent_pop}. The distribution carries a different
interpretation under each route \citep[pp.~52--55,
72--77]{jaynes1979b}[pp.~25--28]{jaynes1982}[\sect~I]{jaynes1982b}{jaynes1986d_r1996}[\sect~11.1]{jaynes1994_r2003}.

\begin{enumerate}[wide,label=(\alph*)]
\item \label{item:entropy_route}One route is the choice of the distribution having the highest
  Shannon entropy, given only a quantitative assessment some of its
  properties, such as expectations. The numerical choice of the value of
  such properties is a (subjective) assumption.

  \medskip

  In the two other routes the maximum-entropy distribution is obtained as
  an \emph{approximation} of a distribution obtained via the probability
  calculus, using data coming from a set of $T$ measurements -- as in our
  present case.\mynote{refs here} Also in this case some (subjective)
  assumptions are necessary: they concern our beliefs about the long-run
  relative frequencies of the measurement outcomes:\footnote{Such
    assumptions are always necessary at the beginning of an inference:
    \enquote{Now the axioms of probability enable us to infer any
      probability-conclusion \emph{only} from probability-premisses. In
      other words, the calculus of probability does not enable us to infer
      any probability-value unless we have some probabilities or
      probability relations \emph{given}. Such data cannot be supplied by
      the mathematician. E.g. the rules of arithmetic and the axioms of the
      probability-calculus are utterly impotent to determine, on the
      supposed knowledge that the throw of a coin must yield either head or
      tail and cannot yield both, the probability that it will yield head
      or that it will yield tail. We must assume that the two co-exclusive
      and co-exhaustive possibilities are \emph{equally probable}, before
      we can estimate the probability of either as being a half of
      certitude} \citep[\emph{Appendix on eduction}, \sect~5,
    p.~182]{johnson1924}.}

\item \label{item:entr_prior_route}In one case we consider all possible
  sets of measurement outcomes to be \emph{roughly} equally likely; this
  leads to a probability for the frequencies $\ynu$ proportional to a
  multinomial coefficient $\binom{L}{L\ynu}$, with $L$ large but smaller
  than $T$. (We cannot assume the sets of measurement outcomes to be
  exactly equally likely, because this is equivalent to their independence:
  \cf\
  \cites[\sect~6.7]{jaynes1994_r2003}[\sect~B]{portamana2009}[\sect~2]{portamana2017}.)
\item \label{item:suff_route}In the other case we assume that the
  measurements have a \emph{sufficient statistics}: the same as appears in
  the exponential of the maximum-entropy distribution.
\end{enumerate}

It's important to keep in mind that the approximate equivalence of these
three routes only holds under very specific assumptions -- which \emph{have
  physical and biological meanings and consequences}. In particular,
route~\ref{item:suff_route} implies that we can discard other empirical
statistics of the data, if they are known; whereas
route~\ref{item:entr_prior_route} requires us to specify all known
empirical statistics, because using only a subset of them may lead to
different results. Route~\ref{item:entropy_route} is also supposed to be
used with all known data. Moreover, the approximate equivalence of
route~\ref{item:entropy_route} with routes~\ref{item:entr_prior_route} and
\ref{item:suff_route} \emph{only holds if $T$ is much larger than the
  possible values of the activity $\yRf$}.
